Iterative Monomorphisation

### 2024-02-14 iterative_monomorphisation.tex
more

  * By a type version of Herbrand's theorem, if a set of polymorphic
    formulas is unsatisfiable, then there exists a finite set of monomorphic
    instances of the formulas that is also unsatisfiable. However, this
    finite set cannot be computed \cite{xxx}; hence any approach based on
    instantiation of type variables is incomplete.

@@

We implemented this algorithm in Zipperpostion, a higher order automatic theorem prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E as a backend. Our algorithm makes it possible, for the first time, to use E with polymorphic problems. Moreover, our implementation of the algorithm in Zipperposition can be used as a preprocessor to interface with other stand-alone provers.

@@

We find that:
%
\begin{enumerate}
\item Zipperposition benefits substantially from the E backend.

@@

\item For Leo-III and Vampire, we find that monomorphisation is indeed more effective than native polymorphism.
\end{enumerate}

>>>>

  * By a type version of compactness, if a set of polymorphic
    formulas is unsatisfiable, then there exists a finite set of monomorphic
    instances of the formulas that is also unsatisfiable. However, this
    finite set cannot be computed \cite{xxx}; hence any approach based on
    instantiation of type variables is incomplete.

@@

We implemented this algorithm in Zipperpostion \cite{xxx}, a higher order automatic theorem prover written in OCaml. Although Zipperposition is polymorphic, it uses the monomorphic prover E \cite{xxx} as a backend. Our algorithm makes it possible, for the first time, to use E with polymorphic problems. Moreover, our implementation of the algorithm in Zipperposition can be used as a preprocessor to interface with other stand-alone provers.

@@

Our findings are as follows:
%
\begin{enumerate}
\item Zipperposition benefits substantially from the E backend.

@@

\item For Leo-III \cite{xxx} and Vampire \cite{xxx}, we find that monomorphisation is indeed more effective than native polymorphism.
\end{enumerate}
